---
title: Using RQuantLib to automatically wrap QuantLib::Date objects
author: Othman El Hammouchi
output: html_document
vignette: >
  %\VignetteIndexEntry{Using RQuantLib to automatically wrap QuantLib::Date objects}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
library(RQuantLibExample)
```

This vignette illustrates how RQuantLib can be used to automatically translate (vectors of) QuantLib::Date objects back and forth between C++ and R.

## Returning a date

`dateInDays` takes in an integer `n` and returns the date `n` days from today.

```{Rcpp, eval=FALSE}
// [[Rcpp::export]]
QuantLib::Date dateInDays(int n) {
    QuantLib::Date today = QuantLib::Date::todaysDate();
    QuantLib::Date res = today + n;
    return res;
}
```

```{r}
Sys.Date()
dateInDays(5)
```

## Accepting dates

`yearFrac` takes in two dates and returns the year fraction between them according to the Act/Act ISDA convention.

```{Rcpp, eval=FALSE}
// [[Rcpp::export]]
double yearFrac(QuantLib::Date date1, QuantLib::Date date2) {
    QuantLib::DayCounter convention =
        QuantLib::ActualActual(QuantLib::ActualActual::ISDA);
    double res = convention.yearFraction(date1, date2);
    return res;
}
```

```{r}
yearFrac(Sys.Date(), as.Date("2024-12-31"))
``` 

## Returning a vector of dates

```{Rcpp, eval=FALSE}
// [[Rcpp::export]]
std::vector<QuantLib::Date> couponDates(QuantLib::Date start, QuantLib::Date maturity, QuantLib::Frequency freq) {
    QuantLib::Schedule schedule =
        QuantLib::MakeSchedule()
            .from(start)
            .to(maturity)
            .withCalendar(QuantLib::TARGET())
            .withConvention(QuantLib::BusinessDayConvention::ModifiedFollowing)
            .withFrequency(freq);
    return schedule.dates();
}
```

```{r}
couponDates(Sys.Date(), as.Date("2024-12-31"), 2)
``` 

## Accepting a vector of dates

```{Rcpp, eval=FALSE}
// [[Rcpp::export]]
std::vector<double> daysBetween(std::vector<QuantLib::Date> dates) {
    std::vector<double> res(dates.size() - 1);
    for (int i = 1; i < dates.size(); i++) {
        res[i - 1] = dates[i] - dates[i - 1];
    }
    return res;
}
```

```{r}
daysBetween(as.Date(c(Sys.Date(), "2024-12-31", "2025-7-17")))
``` 